Hardware-accelerated Rendering of Intersecting Volumes using Boolean Operations

### 2013-04-22 paper.tex
Replaced XML for first approach with pseudocode

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{boolean-and.pdf}
\caption{
Top down view of rendering intersecting volumes using the Boolean AND with depth
buffer masking approach.  First the part of the volume directly behind the
intersection, colored red, is rendered.  Next the intersection, colored green,
is rendered.  Then the the part of the volume directly in front of the
intersection, colored blue, is rendered.  Finally the depth buffer is masked off
using the intersection, and everything else, colored yellow, is rendered.
}
\label{boolean-and}
\end{figure}

@@

\begin{Verbatim}[fontsize=\small]
  <texture id="first-volume-texture" file="first.vlb">
  <texture id="second-volume-texture" file="second.vlb">
  <texture id="coords-texture-1" width="768" height="768">
  <texture id="coords-texture-2" width="768" height="768">
  <texture id="results-texture" width="768" height="768">
\end{Verbatim}

Next the volumes behind the intersection are rendered into the results texture.
Because we want the part of the volume behind the intersection we first render
the texture coordinates of the back faces of the cubes into a framebuffer, then
render the back faces of the intersection as per \cite{Kruger03}.  Note in the code
snippet below we only show the first volume, hide where we set the uniform
variables in the shader program for brevity, and give the intersection an ID so
it can be redrawn later.

\begin{Verbatim}[fontsize=\small]
  <cull faces="front" />
  <framebuffer>
    <attachment usage="color" link="coords-texture-1" />
    <use program="primary-coords-program">
      <instance of="cube-1-group" />
    </use>
  </framebuffer>
  <framebuffer>
    <attachment usage="color" link="results-texture" />
    <use program="first-pass-program">
      <booleanAnd id="intersection" of="cube-1 cube-2" />
    </use>
  </framebuffer>
\end{Verbatim}

@@

\begin{Verbatim}[fontsize=\small]
  <cull faces="front" />
  <framebuffer>
    <attachment usage="color" link="coords-texture-1" />
    <use program="primary-coords-program">
      <instance of="intersection" />
    </use>
  </framebuffer>
  <framebuffer>
    <attachment usage="color" link="coords-texture-2" />
    <use program="secondary-coords-program">
      <instance of="intersection" />
    </use>
  </framebuffer>
  <cull faces="back" />
  <framebuffer>
    <attachment usage="color" link="third-pass-texture" />
    <use program="third-pass-program">
      <uniform name="CoordsTexture1" link="coords-texture-1" />
      <uniform name="CoordsTexture2" link="coords-texture-2" />
      <instance of="intersection" />
    </use>
  </framebuffer>
\end{Verbatim}

After that the parts of the volumes in front of the intersection are rendered to
the results texture.  We have the shader program set the depth of the fragments
generated to the lowest value to effectively mask off that part of the screen.

\begin{Verbatim}[fontsize=\small]
  <cull mode="back" />
  <framebuffer>
    <attachment usage="color" link="coords-texture-1" />
    <use program="primary-coords-program">
      <uniform name="MVPMatrix" usage="modelviewprojection" />
      <instance of="cube-1" />
    </use>
  </framebuffer>
  <framebuffer>
    <attachment usage="color" link="results-texture" />
    <use program="fourth-pass-program">
      <instance of="intersection" />
    </use>
  </framebuffer>
\end{Verbatim}

@@

\begin{Verbatim}[fontsize=\small]
  <blend enabled="true" />
  <depth function="less" />
  <sort>
    <group>
      <framebuffer>
        <attachment usage="color" link="coords-texture-1" />
        <use program="primary-coords-program">
          <cull mode="front" />
          <instance of="cube-1-group" />
        </use>
      </framebuffer>
      <use program="kruger-program">
        <cull faces="back" />
        <group id="cube-1-group">
          <translate x="-1">
            <cube id="cube-1" />
          </translate>
        </group>
      </use>
    </group>
    <group>
      <framebuffer>
        <attachment usage="color" link="coords-texture-2" />
        <use program="primary-coords-program">
          <cull mode="front" />
          <instance of="cube-2-group" />
        </use>
      </framebuffer>
      <use program="kruger-program">
        <cull mode="back" />
        <group id="cube-2-group">
          <translate x="+1">
            <cube id="cube-2" />
          </translate>
        </group>
      </use>
    </group>
  </sort>
\end{Verbatim}

@@

Next we render any pieces of the first cube in front of the intersection to the
results texture.  Similar to the first pass, we use the depth buffer to make
sure we only take the texture coordinates of the faces we want.  Again we omit
the code to render the second cube.

>>>>

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{boolean-and.pdf}
\caption{
Top down view of rendering intersecting volumes using the Boolean AND with depth
buffer masking approach.  First the part of the volume directly behind the
intersection, colored red, is rendered.  Next the intersection, colored green,
is rendered.  Then the the part of the volume directly in front of the
intersection, colored blue, is rendered.  Finally the depth buffer is masked off
using the Boolean AND so that when the cubes are rendered again, only the parts
colored yellow are rendered.
}
\label{boolean-and}
\end{figure}

@@

\begin{enumerate}
  \item Bind volume textures
  \item Bind textures to store texture coordinates
  \item Bind texture to store results
\end{enumerate}

Next the part of the volume behind the intersection is rendered into the results
texture.  To get just that part the back faces of the Boolean AND are used
instead of the front faces of the cube.

\begin{enumerate}
  \item Cull front faces
  \item Bind framebuffer with coordinates texture
  \item Activate shader program for rendering texture coordinates
  \item Draw cube behind intersection
  \item Bind framebuffer with results texture
  \item Activate shader program for rendering a volume
  \item Draw Boolean AND of cubes
\end{enumerate}

@@

\begin{enumerate}
  \item Cull front faces
  \item Bind framebuffer with first coordinates texture
  \item Activate shader program for rendering primary texture coordinates
  \item Draw Boolean AND of cubes
  \item Bind framebuffer with second coordinates texture
  \item Activate shader program for rendering secondary texture coordinates
  \item Draw Boolean AND of cubes
  \item Cull back faces
  \item Bind framebuffer with results texture
  \item Activate shader program for rendering volume
  \item Draw Boolean AND of cubes
\end{enumerate}

After that the part of the volume in front of the intersection is rendered to
the results texture.

\begin{enumerate}
  \item Cull back faces
  \item Bind framebuffer with coordinates texture
  \item Activate shader program for rendering primary texture coordinates
  \item Draw cube in front of intersection
  \item Bind framebuffer with results texture
  \item Activate shader program for rendering volume
  \item Draw Boolean AND of cubes
  \item Mask off depth buffer
\end{enumerate}

The area of the screen with the intersection should be masked off so that
nothing else in that area will be rendered.

\begin{enumerate}
  \item Activate shader program for setting depth to zero
  \item Draw Boolean AND of cubes
\end{enumerate}

@@

\begin{enumerate}
  \item Enable blending
  \item Set depth function to \emph{LESS}
  \item Bind framebuffer with coordinates texture
  \item Cull front faces
  \item Activate shader program for rendering primary texture coordinates
  \item Draw cube behind intersection
  \item Cull back faces
  \item Activate shader program for rendering volumes
  \item Draw cube behind intersection
  \item Repeat for cube in front of intersection
\end{enumerate}

@@

Next we render any pieces of the cube in front of the intersection to the
results texture.  Similar to the first pass, we use the depth buffer to make
sure we only take the texture coordinates of the faces we want.  Again we omit
the code to render the second cube.
